Benefits and risks associated with children’s and adolescents’ interactions with electronic screens: An umbrella review

### 2022-03-08 reports/manuscript_content.Rmd
Finish changes based on comments

Consider children watching television programs---an often cited form of screen time harm.
Indeed, we find robust evidence for a small association with poorer academic performance and literacy skills [@adelantado-renauAssociationScreenMedia2019].
However, we also found evidence that if the content of the program was educational, or the child was watching the program with a parent (i.e., co-viewing), this exposure was instead associated with better literacy [@madiganAssociationsScreenUse2020].
Thus, parents may play an important role in selecting content that is likely to benefit their children or, perhaps, interact with their children in ways that may foster literacy (e.g., asking their children questions about the program).
Similar nuanced findings occur for video games.
The credible evidence we identified showed that video game playing is associated with poorer body composition and learning [@adelantado-renauAssociationScreenMedia2019;@poorolajalBehavioralFactorsInfluencing2020].
However, when the video game was designed specifically to teach numeracy, playing these games showed learning benefits [@byunDigitalGamebasedLearning2018].
One might expect that video games designed to be physically active could confer health benefits, but none of the meta-analyses examining this hypothesis met our thresholds for statistical credibility (see Supplementary Files 4 & 5) therefore this hypothesis could not be addressed.

@@

@@

Screen use research is big, varied, and rapidly growing.
Reviews tended to be general (e.g., all screen time) and even when more targeted (e.g., social media) nuances related to specific content (e.g., Instagram vs Facebook) have not been meta-analysed or have not produced credible evidence. Fewer than 20\% of the effects identified met our criteria for statistical credibility. Most studies which did not meet our critiera did so because they failed to provide study-level data (or, did not provide sufficent study level data such as including effect sizes but not sample sizes), despite <!--TODO: add reference-->

Screen time research has a well-established measurement problem, which impacts the individual studies of this umbrella review.
The vast majority of screen time research relies on self-reported data, which not only lacks the nuance required for understanding the effects of screen time, but may also be inaccurate.
In a meta-analysis of 47 studies comparing self-reported media use with logged measures, Parry et al [@parrySystematicReviewMetaanalysis2021] found that the measures were only moderately correlated ($r = 0.38$), with self-reported problematic usage fairing worse ($r = 0.25$).
It may be that the field of screen time research cannot be sufficiently advanced until more accurate and nuanced measures are available.

>>>>

Consider children watching television programs---an often cited form of screen time harm.
We found robust evidence for a small association with poorer academic performance and literacy skills [@adelantado-renauAssociationScreenMedia2019].
However, we also found evidence that if the content of the program was educational, or the child was watching the program with a parent (i.e., co-viewing), this exposure was instead associated with better literacy [@madiganAssociationsScreenUse2020].
Thus, parents may play an important role in selecting content that is likely to benefit their children or, perhaps, interact with their children in ways that may foster literacy (e.g., asking their children questions about the program).
Similar nuanced findings occur for video games.
The credible evidence we identified showed that video game playing is associated with poorer body composition and learning [@adelantado-renauAssociationScreenMedia2019;@poorolajalBehavioralFactorsInfluencing2020].
However, when the video game was designed specifically to teach numeracy, playing these games showed learning benefits [@byunDigitalGamebasedLearning2018].
One might expect that video games designed to be physically active could confer health benefits, but none of the meta-analyses examining this hypothesis met our thresholds for statistical credibility (see Supplementary Files 4 & 5) therefore this hypothesis could not be addressed.

@@

Among studies that met our criteria for statistical certainty heterogeneity was high, with almost all effects having $I^2 > 50\%$.
Much of this heterogeneity is likely explained by differences in measures across pooled studies, or in some cases, the generic nature of some of the exposures.
For example, "TV programs and movies" covers a substantial range of content, which may explain the heterogeneous association with education outcomes.

@@

Screen use research is big, varied, and rapidly growing.
Reviews tended to be general (e.g., all screen time) and even when more targeted (e.g., social media) nuances related to specific content (e.g., Instagram vs Facebook) have not been meta-analysed or have not produced credible evidence.
Fewer than 20\% of the effects identified met our criteria for statistical credibility.
Most studies which did not meet our critiera failed to provide study-level data (or did not provide sufficent data, such as including effect estimates but not sample sizes).
Newer reviews were more likely to provide this information than older reviews, but it highlights the importance of data and code sharing as recommended in the PRISMA guidelines [@pagePRISMA2020Statement2020].
When study level data was available, many effects were removed because the pooled sample size was small, or because there were fewer than ten studies on which to perform an Egger's test.
It seems that much of the current screen time research is small in scale, and there is a need for larger, high-quality studies.

Screen time research has a well-established measurement problem, which impacts the individual studies of this umbrella review.
The vast majority of screen time research relies on self-reported data, which not only lacks the nuance required for understanding the effects of screen time, but may also be inaccurate.
In a meta-analysis of 47 studies comparing self-reported media use with logged measures, Parry et al [@parrySystematicReviewMetaanalysis2021] found that the measures were only moderately correlated ($r = 0.38$), with self-reported problematic usage fairing worse ($r = 0.25$).
Indeed, of 622 studies which measured the screen time of 0---6 year-olds, only 69 provided any sort of psychometric properties for their measure, with only 19 studies reporting validity [@byrneMeasurementScreenTime2021].
While some researchers have started using newer methods of capturing screen behaviours---such as wearable cameras [@smithFeasibilityAutomatedCameras2019] or device-based loggers [@rydingPassiveObjectiveMeasures2020]---these are still not widely used.
It may be that the field of screen time research cannot be sufficiently advanced until accurate, validated, and nuanced measures are more widely available and adopted.
